Security vs. Maintainability: \\ Fixing Vulnerabilities Obfuscates your Code

### 2019-01-29 paper/ICPC19/icpc19.tex
minor fixes

\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage[table,xcdraw]{xcolor}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\definecolor{mypink3}{cmyk}{0, 0.7808, 0.4429, 0.1412}

@@

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by OpenSSL developers to the
CVE-2014-3506 vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list) /*!\annotation{lst:func2}!*/
+     return s->max_cert_list;
+  return max_len;
+ }

@@

As saids before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintanability reports of the refactorings of
each project. Figure \ref{fig:guidelines} presents the 10 guidelines
proposed by BCH's authors for delivering software that is easy to maintain
\cite{Visser:2016:OREILLY}.

\begin{figure}[H]
 	\centering 	\includegraphics[width=0.49\textwidth]{figures/guidelines.pdf}
 	\caption{Guidelines to produce maintainable code}
	\label{fig:guidelines}
\end{figure}

@@

\emph{Units} are the smallest groups of code that can be maintained and executed
independently \cite{Visser:2016:OREILLY}. McCabe \cite{1702388} is a software
metric that calculates the number of linearly independent paths of the source
code. BCH uses this metric to measure the number of branch points in the
\emph{Write Simple Units of Code} guideline. The bar represents the top 30 of
units that violates the guideline, sorted by severity, which is indicated by the
colors of the checkboxes. The green bar represents the number of compliant
branch points per unit (\emph{at most 5}), i.e., the units are compliant with
ISO 25010 \cite{iso:2011}. Yellow, orange and red bars represent units that do
not comply with medium (\emph{above 5}), high (\emph{above 10}) and very high
(\emph{above 25}) severity levels. In the bar, there are marks that pinpoint the
compliance thresholds for each severity level.

@@

\noindent
where $G$ is the group of selected maintainability guidelines from BCH (Figure
\ref{fig:guidelines}) and $v$ is the version of the software under evaluation.
The maintenance $M$ based on the guideline $g$ for a given version of a project
is computed with the following equation:

@@

Projects with Broken Authentication flaws are the most affected ones by security refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any change. The results also show that fixing memory leaks affects the maintainability of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected. 35\% (28) are affected positively. Refactoring vulnerabilities that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of the cases exhibit maintainability improvement while 22.5\% (9) show that overall maintainability does not suffer any change. Miscellaneous refactorings pattern integrates several types of security flaws whose impact is of 43.1\% (84 cases) negative in the maintainability of open-source sofware. 31.8\% (62) of the cases do not exhibit any change in maintainability and 25.1\% (49) of cases have a positive impact on the maintainability of open-source software.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub. 
Since the dataset has a broad group of open-source projects with different backgrounds, this metric may not be suitable for all of them. However, BCH does have a representative benchmark of
closed and open-source software projects that is under the calculation of the thresholds used in the computation of each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012} (Figure \ref{fig:guidelines}).

@@

>>>>

\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage[table,xcdraw]{xcolor}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}

@@

\usepackage{balance}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by OpenSSL developers to the
\\CVE-2014-3506 vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list) /*!\annotation{lst:func2}!*/
+     return s->max_cert_list;
+  return max_len;
+ }

@@

As said before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintanability reports of the refactorings of
each project. Table \ref{tab:guidelines} presents the 10 guidelines
proposed by BCH's authors for delivering software that is easy to maintain
\cite{Visser:2016:OREILLY}.

\begin{table}[H]
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{2.5cm}L{5.5cm}}

\toprule
\textbf{10 Guidelines} & \textbf{Description}\\
\midrule
\textbf{Write Short Units of Code} & Limit code units to 15 LOCs because smaller units are easier to understand, reuse and test them\\\midrule
\textbf{Write Simple Units of Code} & Limit branch points to 4 per unit because it makes units easier to test and modify \\\midrule
\textbf{Write Code Once} & Do not copy code because bugs tend to replicate at multiple places (inefficient and error-prone)\\\midrule
\textbf{Keep Unit Interfaces Small} & Limit the number of parameters to at most 4 because it makes units easier to understand and reuse\\\midrule
\textbf{Separate Concerns in Modules} & Avoid large modules because changes in loosely coupled databases are easier to oversee and execute\\\midrule
\textbf{Couple Architecture Components Loosely} & Minimize the amount of code within modules that is exposed to modules in other components\\\midrule
\textbf{Keep Architecture Components Balanced} & Balancing the number of components ease locating code and allow for isolated maintenance\\\midrule
\textbf{Keep your Codebase Small} & Reduce and avoid the system size because small products are easier to manage and maintain\\\midrule
\textbf{Automate Tests} & Test your codebase because it makes development predictable and less risky\\\midrule
\textbf{Write Clean Code} & Avoid producing software with code smells because it is more likely to be maintainable in the future\\
\bottomrule
\end{tabular}
\label{tab:guidelines}
\end{table}

@@

\emph{Units} are the smallest groups of code that can be maintained and executed
independently \cite{Visser:2016:OREILLY}. McCabe \cite{1702388} is a software
metric that calculates the number of linearly independent paths of the source
code. BCH uses this metric to measure the number of branch points in the
\emph{Write Simple Units of Code} guideline. The bar represents the top 30 of
units that violates the guideline, sorted by severity, which is indicated by the
colors of the checkboxes. The green bar represents the number of compliant
branch points per unit (\emph{at most 5}), i.e., the units are compliant with
ISO 25010 \cite{iso:2011}. Yellow, orange and red bars represent units that do
not comply with medium (\emph{above 5}), high (\emph{above 10}) and very high
(\emph{above 25}) severity levels. In the bar, there are marks that pinpoint the
compliance thresholds for each severity level.
\Sof{Adress thresholds confusion}

@@

\noindent
where $G$ is the group of selected maintainability guidelines from BCH (Table
\ref{tab:guidelines}) and $v$ is the version of the software under evaluation.
The maintenance $M$ based on the guideline $g$ for a given version of a project
is computed with the following equation:

@@

Projects with Broken Authentication flaws are the most affected ones by security refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any change. The results also show that fixing memory leaks affects the maintainability of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected. 35\% (28) are affected positively. Refactoring vulnerabilities that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of the cases exhibit maintainability improvement while 22.5\% (9) show that overall maintainability does not suffer any change. Miscellaneous refactorings pattern integrates several types of security flaws whose impact is of 43.1\% (84) negative in the maintainability of open-source sofware. 31.8\% (62) of the cases do not exhibit any change in maintainability and 25.1\% (49) of cases have a positive impact on the maintainability of open-source software.

@@

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub.
Since the dataset has a broad group of open-source projects with different backgrounds, this metric may not be suitable for all of them. However, BCH does have a representative benchmark of
closed and open-source software projects that is under the calculation of the thresholds used in the computation of each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012} (Table \ref{tab:guidelines}).

@@

\balance
